---
title: "Looking at the Data"
author: "Camille Vasquez"
date: '2022-08-08'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
## Case Study Questions:
* What are some trends in smart device usage?
* How could these trends apply to Bellabeat customers?
* How could these trends help influence Bellabeat marketing strategy?


## How many different Participates in each set?
What Variables are we working with in this project?
```{r}
cat("Partipates for Daily Activity:", n_distinct(daily_activity$Id),"\n")
cat("Partipates for Daily Steps:", n_distinct(daily_steps$Id),"\n") 
cat("Partipates for Daily Calories", n_distinct(daily_calories$Id),"\n") 
cat("Partipates for Daily Intensities:",n_distinct(daily_intensities$Id),"\n") 
```

## Looking at the dataframe Daily Activity

```{r}
daily_activity %>%  
  select(TotalSteps,
         TotalDistance,
         SedentaryMinutes) %>%
  summary()
```
Looking at the user's average, we can see the realtionship between total steps, 
total distance, and sedentary minutes. This will give us an idea of what our 
main focus should be on if we wish to find trends in the data and use the trends
to putt more users in. 
## Searching for Trends in the Dataframe "Daily Steps"
Since we are looking for trends in the data lets look into the average for 
each user.
Using the dataframe daily_steps, the graph below demonstrates the usage of the 
device depending on the day. With the data in the first graph many of the users 
fall below the 10,000 step mark. Since in fitness, the goal is to make at least 
10,000 steps a day, it would be a ideal marking strategy to either remind the 
users or have the users set a goal for themselves for motivation. 

```{r}
daily_steps %>%  
  select(ActivityDay,
         StepTotal) %>%
  summary()
```

We can see from the summary of our data frame that our mean is an average of 
7,638 steps a day. Which falls below the recommended average.
We may also varify by looking at a visual represenation of

```{r}
#ggplot(data=daily_steps, aes(x=ActivityDay,y=StepTotal))+ geom_point() 
#+ geom_point(aes(x=10000))
#ggplot(data=daily_steps, aes(x=ActivityDay,y=StepTotal))+ geom_point() 
#+ geom_smooth(method="gam", 
#formula = y ~s(x))+facet_wrap(~Id)+labs(title="Activity Day vs Steps Total",
#       caption=paste0("Data from: April 2016", " to May 2016"),
#       x="Activity Day",
#       y="Steps Total")
daily_steps%>%
  group_by(Id)%>%
  #mutate(Id = fct_reorder(Id, 1:33)) %>%
  ggplot(.,aes(x = Id,y = StepTotal)) + geom_bar(stat = "identity")  
#+ scale_x_discrete(limits=daily_steps$Id)
```
For this graph, I know I need to fix the columns for "Id" and change them to 
numeral values (values of the rows) or change the axis. 
Otherwise, we can see the total steps of our users above. 

## Comparing Sedentary Minutes to Total Steps
Sedentary minutes refers to the day's non-excerises like siting down or standing 
up once a while. We want to see if there is a trend between in the activty. 

```{r}
ggplot(data=daily_activity, aes(x=SedentaryMinutes, y=TotalSteps)) + geom_point() + geom_smooth(method="gam", formula = y ~s(x))
```
Using the graph above, we can see that the trend is a bit harder to see but the 
users who send more time sitting do not accomplish their 10,000 goal. 

## Total Steps verus Total Calories 
Let us look at the relationship between total steps taken and calories burned 
througout the day. 
```{r}
ggplot(data=calories_and_steps, aes(x=StepTotal, y=Calories)) + geom_point() 
+ geom_smooth(method="gam", formula = y ~s(x))
```
Clearly we see a relationship between steps taken and calories in the day. 
We can use this to our advantage and explain to our stakeholders how we can 
market towards new users.

## Conclusion 
We have been able to show some relationships in the data to come to a 
data-driven-decision on how the marketing team should plan their next approach. 
It may be beneficial to market towards customers looking to track their steps
throughout the day and show how it can lead to burning more calories. Another 
idea is how the user who more than likely have office jobs and excessive amount 
of daily sedentary minutes are still able to reach a 10,000 steps goal.
